The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-07-25 paper/referee/response-revision1.txt

MN-14-1011-MJ: Authors' response to the reviewer's comments
===========================================================

We are grateful to the referee for the insightful feedback, which has helped us
to improve the presentation of the data products.  In his/her report, the 
referee identified areas where additional information would (i) strengthen the 
role of the paper as a user's guide to the survey, and (ii) would lessen the 
burden on the reader to track down information from previous works.  
We appreciate the merits of the referee's suggestions in this respect, and have 
acted upon them by adding a total of N new pages, 1 new table, and 1 new figure 
to the manuscript.  These changes have been marked in bold in the manuscript.  
We have also added N new sections of auxillary material to our web site as a 
direct consequence.

In what follows we will start by explaining our response to each of the issues 
which the referee identified as high-level remarks (9 items), followed by our 
response to the more specific remarks (8 items).

High-level items
----------------

[Referee]
> 1. This manuscript presents a large amount of information about IPHAS and the preparation
> of and contents of the data products.  It also relies heavily on previous
> publications for general descriptions of the survey, particulars of the instrumentation,
> data processing algorithms and earlier data releases.   As such, it puts considerable burden 
> on the readers to track down information from many different sources.  It is
> understandable that the authors prefer to not duplicate information that is already
> available elsewhere, but it will make it much easier to understand and use the IPHAS data
> products if at least some of the general information  could be repeated in this
> manuscript.
>
> For example, it would be helpful to include concise summary of the IPHAS survey
> parameters and this data release in tabular form.  This summary would serve as
> a "IPHAS-At-A-Glance" for users, and coule include the survey area, dates, telescope,
> camera and survey bandpasses, etc.  Much of the information is presented in the paper,
> but widely spread over the manuscript.  Without at least a guide to finding the
> information, users will be challenged to find what they need.

[Authors' response]
We have kept the referee's recommendation to lessen the burden on the reader in
mind while revising the paper in response to the items to follow below.
As request by the referee, we have included a concise summary table of the
survey parameters at the start of Section 2, in which the 15 key properties
of the survey are summarised (Table 1).

[Referee]
> 2. It was not immediately obvious if the images and catalogue from the Second IPHAS
> release supersede those in the 2008 Initial Data Release.  I believe the
> Second Release includes a reprocessing of all of the survey data to-date, so it
> is intended to replace the Initial Release, but that should be stated explicitly
> in the Introduction (e.g. in the 3rd paragraph).
>
> If my assumption is correct, there should be a summary statement about why the
> Second Release images and catalogue are in fact better products.  For example,
> better calibrations, processing algorithms, etc.

[Authors' response]
The referee correctly identified that we forgot to state that this new release 
supersedes the previous one.  We have added an explicit statement in this sense
to the third paragraph of Section 1, and in addition we have edited the last
paragraph of Section 3.1 to list the key differences between the past and the
present release.

[Referee]
> 3. As stated in Section 7.2, the aim of IPHAS is to enable the discovery of
> emission line objects in the Galactic Plane.  Consequently, I was very surprised
> that no discussion of the sensitivity to line emission (e.g. H-alpha EW) was
> presented in the manuscript.  Perhaps this will be addressed further in the planned
> Barentsen et al paper, but it seems that at least some rudimentary demonstration
> of the sensitivity to line emission is a requirement for this paper.  The
> general photometric sensitivity in the three bands is discussed in section 6.2, and
> an illustration is given how to find emission-line candidates using the color-color
> diagram in Figure 18.  I strongly recommend that at least some examples be given
> of the r-H-alpha color excess as a function of line EW for a sample of objects
> with published spectra.

[Authors' response]
The referee requested a discussion on the sensitivity of the H-alpha 
photometry to line emission to be included in this work.  In response, we have
added one new section (Sect 7.3) and one new figure (Fig 19) to the paper.
In this new section, which is titled "r-Halpha excess as a quantitative measure
of Halpha emission", we compare the spectroscopic H-alpha EWs for a sample of
classical Be stars from Raddi et al (2014) against the position of these
objects in the IPHAS colour-colour diagram.  A good agreement is found.
Moreover, the accompanying discussion offers a brief guide to the previous 
works on this topic.

[Referee]
> 4. Many of the successful large surveys include a list of known caveats concerning
> released data products to help user's avoid pitfalls in their research.  I suggest that
> the authors generate a short list of the most important things about which users
> should be aware when using IPHAS data.  Users appreciate knowing how to avoid mistakes
> due to misinterpretation of known issues, and do not think less of a product because
> of them.

[Authors' response]
In response to the referee's suggestion, we have (i) introduced a summary of
the most common pitfalls at the start of Section 6, and (ii) have added a
caveats section to our data release web page (www.iphas.org/dr2).
We note that it is not possible, at this stage, to include an exhaustive list
of all the caveats, because we cannot anticipate all forms of use of the data.
Hence in Section 6 we alert the reader to the fact that we offer a caveats
section on the website, which will be updated as user experience accumulates.

[Referee]
> 5. In general, the images do not receive the same level of description in the manuscript
> as does the Catalogue.  Any guidance that can be given to users about them would be
> beneficial.  For example, at the very least include a description of the FITS image
> headers, analogous to the descriptions of the catalog columns that is given in
> the Appendix.  They should also mention that tables containing the image metadata are available at IPHAS web site.

[Authors' response]
Although the paper intended to focus on the catalogue product, we agree with the
referee that it would be beneficial to give users additional guidance on the use 
of the image products.  For this reason, we have made significant improvements 
to our web site by (i) providing a query tool to retrieve IPHAS images by 
coordinates, and (ii) provide more comprehensive documentation on the metadata, 
the FITS header keywords, and the flux calibration (see www.iphas.org/images).
Section 6 of the paper has been edited to advertise these resources, which are 
set to be updated in the future whilst we prepare a data release which focuses
more strongly on image and image mosaic products.

[Referee]
> 6. The fact that only 39% of the catalogue entries satisfy the criteria necessary to be
> classified as "reliable" begs the question "just what are the other 61% of objects."
> The authors should add more information about the general reliability of the
> catalogue with and without their criteria. For example, If a users selects
> "reliable" category sources, what should they expect the net reliability to be?
> Certainly not 100%.  How does reliability vary with source magnitude, SNR, or other 
> measurement parameters?  How much does the reliability vary with source environment
> (e.g. crowding, complex backgrounds)?
>
> Reliability can be gauged in a number of ways, such as comparisons with external,
> deeper catalogues (e.g. SDSS, UKIDSS GPS).

[Authors' response]
The referee questioned the nature of the objects which are marked as "reliable"
in the catalogue, and in particular questioned the nature of the large number of 
sources not flagged as "reliable".  Having considered this comment, we have come 
to conclude that we made a poor choice of terminology by adopting the labels 
"reliable" and "veryReliable" as quality qualifiers.
These classes, which are based on stringent quality criteria set out in 
Tables 6 and 7, aim to help users identify high-quality objects, e.g. by 
demanding a 10-sigma detection in all bands.  It goes without saying that 
there are plenty of objects which do not strictly satisfy these criteria but are 
nevertheless genuine astrophysical objects.  The qualifier "reliable" might 
however confer the incorrect impression that such sources are "unreliable".
As a result, we have decided to change the terminology by renaming the class 
previously known as "reliable" into "a10" (which is short for "all-band 10-sigma"), 
and renaming "veryReliable" into "a10point".  We believe this change in 
terminology will better serve as an operational description of the detection quality.

In addition to this change in terminology, we have followed the referee's suggestion
to gauge the reliability of our objects against an external catalogue.  That is,
we have introduced a new Section 6.2, titled "Source reliability", in which the 
reliability of IPHAS sources as a function of the quality criteria is gauged 
by comparing out catalogue against SDSS.  Moreover, we note that a separate paper 
is in an advanced state of preparation (Farnhill et al.), which will investigate 
the issues surrounding the completeness and reliability of our catalogue, 
e.g. by injecting artificial sources into the images and measuring their recovery 
rates, whilst also comparing source counts between SDSS and IPHAS in more detail.

[Referee]
> 7. The authors have done a good job describing the photometric measurements and
> calibration accuracy.  However, there is relatively little discussion of the
> astrometric position reconstruction other than saying it is done relative to
> the 2MASS Point Source Catalogue, and there is no validation of the astrometric
> quality of the Catalogue other than to mention matching success.  Some quantification
> of the astrometric performance should be added (e.g. mean position residuals as a function
> of magnitude from comparisons with the astrometric reference stars, or external
> astrometric collections, such as UCAC4).
>
> Positions in the IPHAS Catalogue have been reconstructed with respect to 2MASS.
> However, the authors do not state if the proper motion of the 2MASS reference stars 
> was taken into account.  Although small, the net motion of the reference stars can
> collectively imprint on the Catalogue because the IPHAS observations were made at
> different epochs than 2MASS, in some cases as much as 12 years later.   
> Failure to do so can imprint the proper motion onto IPHAS, and also because the 
> IPHAS measurements span nearly 10 years (2003-2012), there can be secular variations.

[Authors' response]
astrometry ...
proper motions ...

(There does not appear to be an item number 8 in the reviewer's comments.)

[Referee]
> 9. The authors point out in the introduction that IPHAS will complement a number of
> new infrared surveys of the Galactic Plane.  The paper would benefit from at least
> one cross-comparison with the catalogue of one of those surveys, such as GLIMPSE
> or MIPSGAL, as an illustration of how the combination of surveys will enable
> new science.  At present, the only tie-in to the infrared surveys is the
> visual comparison between IPHAS and GLIMPSE images of one field.

[Authors' response]
In response to the referee's suggestion to include a cross-comparison between
IPHAS and an infrared survey, we have added a brief discussion on the overlap
between the GLIMPSE and IPHAS catalogues at the end of Section 7.2.  The 
discussion warns the reader that an optical survey will naturally be 
extinction-limited.

### 2014-07-25 paper/iphas-dr2-rev1.tex
Edits for rev1

\subsection{Reliability}
\label{sec:reliability}

@@

\begin{table}
    \caption{Reliability values as computed by crossmatching the point 
             source catalogues of IPHAS and SDSS DR7 for a number of 
             subsets of IPHAS DR2 across 1 deg$^2$ regions.}
    \label{tbl:reliability}
    \begin{center}
        \begin{tabular}{lcccccc}
        \toprule
        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{a10} & \emph{a10point} \\
        \midrule
        (149.39, 4.06) & 92.95 & 98.83 & 98.53 & 98.54 & 99.67 & 99.82 \\
        (186.59, -2.50) & 89.99 & 98.82 & 98.34 & 98.44 & 99.36 & 99.57 \\
        (202.70, -1.75) & 89.42 & 98.46 & 97.8 & 98.11 & 99.4 & 99.65 \\
        \bottomrule
        \end{tabular}
    \end{center}
\end{table}

@@

\begin{figure}
\captionsetup[subfigure]{labelformat=empty}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability.pdf}
\caption{}
\label{fig:pairmag}
\end{subfigure}
\hspace{0.5cm}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-a10point.pdf}
\caption{}
\label{fig:pairmag_reliable}
\end{subfigure}
\caption{Photometric repeatability as a function of magnitude
         for all sources in the catalogue (panel a)
         and for the \emph{a10point} sources alone (panel b).
         Black dots show the mean absolute residuals
         between the primary and the secondary detections.
         The error-bars show the standard deviations.
         The red trend line shows the average Poissonian uncertainties
         we derived in Fig.~\ref{fig:uncertainties}.
         We find that the \emph{a10point} quality criteria are successful
         at removing objects with large residuals.
    }
\end{figure}
\begin{figure*}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
    \caption{Number of entries in the IPHAS DR2 source catalogue
    as a function of Galactic longitude.
    The upper blue line shows the number of sources
    counted in 1\degr-wide longitude bins.
    The lower red line uses the same binning
    but includes only those sources 
    for which the \emph{deblend} flag is {\sc false}, 
    i.e. unconfused sources for which the CASU pipeline
    did not have to apply a deblending procedure.
    In both cases we counted only those sources
    in the latitude range $-5\degr<b<+5\degr$,
    such that one may obtain a rough guide to source density
    by dividing the counts by 10~deg$^2$.
    The global variations in the source counts
    traces the structure of the Galaxy
    and the distribution of foreground extinction,
    but is also affected by instrumental effects
	such as variations in the observed depth
	and completeness (see text).
   }
    \label{fig:sourcecount}
\end{figure*}

@@

Sh 2-82 is one of a large population of poorly-studied star-forming regions
located at low Galactic latitudes,
which have only recently started to become revealed
by efforts to catalogue the wealth of `bubbles' detected
at mid-infrared wavelengths \citep{Churchwell2006,Simpson2012},
and by efforts to catalogue previously unknown clusters seen 
in the near-infrared \cite[e.g.][]{Bica2003}.
IPHAS data can offer a handle
on the extinction, distance
and stellar contents of these unexplored regions.

{\bf Moreover, IPHAS provides the much-needed counterpart
to the source catalogues being provided by the recent infrared surveys.
For example, amongst the 1810 high-quality sources considered in Fig.~\ref{fig:emitters},
1623 objects (90\%) have a counterpart in the Spitzer catalogue,
and 371 objects (20\%) have a counterpart in the 
Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
both within a matching radius of 2 arcsec.}

%TODO: This is an example of a high-extinction sightline where
% optical observations will naturally be extinction-limited.
% IPHAS will naturally count far fewer sources than (N)IR surveys.
% In the outer-plane the situation is, in fact, reversed,
% except where there is UKIDSS GPS coverage (i.e. dec > 60).
% For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...

@@

The catalogue will be made available through the
Vizier service (http://vizier.u-strasbg.fr),
where it can be queried using a web interface
and using Virtual Observatory (VO) protocols.
In addition, the catalogue can be downloaded
in its entirety from our website (www.iphas.org)
as a collection of binary FITS tables,
each covering a $5\degr\times5\degr$ tile
of the footprint and comprising 50~gigabyte
in total.

We do not recommend using the catalogue
to study extended objects which
are larger than the
aperture diameters specified in this work.
To enable the analysis of diffuse sources,
our website provides access to the pipeline-processed
imaging data. These have been updated to include
a new FITS header keyword, called {\sc photzp},
which contains the re-calibrated zeropoint.
This keyword can be used to convert the number counts $DN$,
i.e. the pixel values in the images,
into Vega-based magnitudes $m$ using:
\begin{equation}
\begin{split}
   m  = \textsc{photzp} - 2.5 \log_{10}(DN).
\label{eqn:mag}
\end{split}
\end{equation}
The {\sc photzp} value has been computed such that it
absorbs the required corrections for atmospheric extinction,
gain variations, exposure time, and the re-calibration shift.
As these images still include moonlight and other sources
of non-astronomical background, they can only support flux measurements
that include a suitably-chosen local background subtraction.

@@

The image repository on our website
is accompanied by an archive of detection tables
-- one for each image --
which may be used to retrieve secondary detections that are not included
in the DR2 source catalogue.
We warn that this set of original images and detection tables
includes data that did not pass quality control,
and has not been globally re-calibrated.
Such data are flagged and must be used with great caution.

>>>>

\subsection{Source reliability}
\label{sec:reliability}

@@

\begin{table}
    \caption{Fractions of sources in our catalogue
             which are also found in the SDSS DR7 catalogue
             within a cross-matching radius of 0.5~arcsec.
             These statistics can be interpreted
             as lower limits
             for the reliability of our sources,
             i.e. the probability that an entry in our
             catalogue is a genuine astrophysical source
             rather than an instrumental artefact.
             The recovery fractions were computed
             in three 1 deg$^2$ circular areas
             where both surveys overlapped.
             The recovery rates are shown
             for IPHAS sources 
             without selection criteria applied (``all''),
             using only objects detected in all three bands (``$nBands=3$''),
             using objects detected at 10$\sigma$
             in all three bands (``a10''),
             and finally using the detections
             which we classified earlier
             as reliable 10$\sigma$ point sources (``a10point'').
    		%Reliability values as computed by crossmatching the point 
             %source catalogues of IPHAS and SDSS DR7 for a number of 
             %subsets of IPHAS DR2 across 1 deg$^2$ regions.
             }
    \label{tbl:reliability}
    \begin{center}
        \begin{tabular}{lcccccc}
        \toprule
        Region ($\ell$, $b$) & all & $nBands=3$ & \emph{a10} & \emph{a10point} \\
        \midrule
        (149.39, 4.06) & 93.0\% & 98.8\% & 99.7\% & 99.8\% \\
        (186.59, -2.50) & 90.0\% & 98.8\% & 99.4\% & 99.6\% \\
        (202.70, -1.75) & 89.4\% & 98.5\% &99.4\% & 99.7\% \\
        \bottomrule
        \end{tabular}
    \end{center}
\end{table}

@@

\begin{figure}
\captionsetup[subfigure]{labelformat=empty}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability.pdf}
\caption{}
\label{fig:pairmag}
\end{subfigure}
\hspace{0.5cm}
\begin{subfigure}[b]{\linewidth}
\centering
\includegraphics[width=0.9\textwidth]{figures/repeatability/repeatability-a10point.pdf}
\caption{}
\label{fig:pairmag_a10}
\end{subfigure}
\caption{Photometric repeatability as a function of magnitude
         for all sources in the catalogue (panel a)
         and for the \emph{a10point} sources alone (panel b).
         Black dots show the mean absolute residuals
         between the primary and the secondary detections.
         The error-bars show the standard deviations.
         The red trend line shows the average Poissonian uncertainties
         we derived in Fig.~\ref{fig:uncertainties}.
         We find that the \emph{a10point} quality criteria are successful
         at removing objects with large residuals.
    }
\end{figure}
\begin{figure*}
    \vspace{1cm}
    \includegraphics[width=\textwidth]{figures/sourcecount/sourcecount.pdf} 
    \caption{Number of entries in the IPHAS DR2 source catalogue
    as a function of Galactic longitude.
    The upper blue line shows the number of sources
    counted in 1\degr-wide longitude bins.
    The lower red line uses the same binning
    but includes only those sources 
    for which the \emph{deblend} flag is {\sc false}, 
    i.e. unconfused sources for which the CASU pipeline
    did not have to apply a deblending procedure.
    In both cases we counted only those sources
    in the latitude range $-5\degr<b<+5\degr$,
    such that one may obtain a rough guide to source density
    by dividing the counts by 10~deg$^2$.
    The global variations in the source counts
    traces the structure of the Galaxy
    and the distribution of foreground extinction,
    but is also affected by instrumental effects
	such as variations in the observed depth
	and completeness (see text).
   }
    \label{fig:sourcecount}
\end{figure*}

@@

Sh 2-82 is one of a large population 
of poorly-studied star-forming regions
located at low Galactic latitudes,
which have only recently started to become revealed
by efforts to catalogue the wealth of `bubbles' detected
at mid-infrared wavelengths \citep{Churchwell2006,Simpson2012},
and by efforts to catalogue previously unknown clusters seen 
in the near-infrared \cite[e.g.][]{Bica2003}.
IPHAS data can offer a handle
on the extinction, distance
and stellar contents of many of such unexplored regions.
{\bf For example,
in the region studied above,
the IPHAS catalogue and the Spitzer/GLIMPSE archive
have 4798 sources in common
within a cross-matching radius of 1 arcsec,
(out of a total of 10\,739 and 11\,321 entries in the
IPHAS and Spitzer catalogues, respectively).
This is a favourable statistic,
given that this is a high-extinction sightline towards the inner plane
where an optical survey will naturally be extinction limited,
and IPHAS can be shown not to reach beyond 4-5\,kpc \citep{Sale2014}.
The situation will be even more favourable
in the outer-plane, where dust optical depths are low.

%where IPHAS is not limited by extinction
%This is an example of a high-extinction sightline where
%optical observations will naturally be extinction-limited.
%IPHAS will naturally count far fewer sources than (N)IR surveys.
%In the outer-plane the situation is, in fact, reversed,
%except where there is UKIDSS GPS coverage (i.e. dec > 60).
%For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...

%Amongst the 1810 high-quality sources which we
%plotted in Fig.~\ref{fig:emitters},
%no less than 1561 objects (87\%) have a counterpart
%in the Spitzer GLIMPSE archive
%within a matching radius of 1 arcsec.
%In the other direction, out of the 6753 objects
%which are in the Spitzer/GLIMPSE catalogue for this region,
%2984 (44\%) have a counterpart in IPHAS.

% are 6753 region considered in Fig.~\ref{fig:emitters}
%contains 6753 objects in the
%Spitzer GLIMPSE catalogue

%371 objects (20\%) also have a counterpart in the 
%Wide-field Infrared Survey Explorer catalogue (WISE; Cutri et al. 2013),
%both within a matching radius of 2 arcsec.

% reliable & r < 20 & (l > 53.55-10/60.) & (l < 53.55+10/60.)                  & (b > 0.01-7.5/60.) & (b < 0.01+7.5/60.)
% Vizier II/293
%TODO: This is an example of a high-extinction sightline where
% optical observations will naturally be extinction-limited.
% IPHAS will naturally count far fewer sources than (N)IR surveys.
% In the outer-plane the situation is, in fact, reversed,
% except where there is UKIDSS GPS coverage (i.e. dec > 60).
% For example, in the anti-centre region where the dust optical depths are low, IPHAS contains more...

@@

The catalogue will be made available through the
Vizier service (http://vizier.u-strasbg.fr),
where it can be queried using a web interface
and using Virtual Observatory (VO) protocols.
In addition, the catalogue can be downloaded
in its entirety from our website
as a collection of binary FITS tables,
each covering a $5\degr\times5\degr$ tile
of the footprint and comprising 50~gigabyte
in total (see www.iphas.org/dr2).

We do not recommend using the catalogue
to study extended objects which
are larger than the
aperture diameters specified in this work.
{\bf To enable the analysis of diffuse sources,
our website provides access to the entire set
of pipeline-processed imaging data,
the associated metadata,
and the documentation describing the FITS header keywords
(see www.iphas.org/images)}.
The image headers have been updated to include
a new keyword, called {\sc photzp},
which contains the re-calibrated zeropoint.
This keyword can be used to convert the number counts $DN$,
i.e. the pixel values in the images,
into Vega-based magnitudes $m$ using:
\begin{equation}
\begin{split}
   m  = \textsc{photzp} - 2.5 \log_{10}(DN).
\label{eqn:mag}
\end{split}
\end{equation}
The {\sc photzp} value has been computed such that it
absorbs the required corrections for atmospheric extinction,
gain variations, exposure time, and the re-calibration shift.
As these images still include moonlight and other sources
of non-astronomical background, they can only support flux measurements
that include a suitably-chosen local background subtraction.

@@

We warn that the image repository on our website
includes data that did not pass quality control
and has not been globally re-calibrated.
Such data are flagged {\bf in the metadata table}
and must be used with great caution.
